A linearization procedure for constrained multibody systems

### 2012-04-05 PetersonGedeHubbard2012.tex
Minor edits.

\begin{abstract}
  We present a systematic procedure for linearizing equations of motion
  governing constrained multibody systems which have been derived with Kane's
  method.  The procedure avoids differential algebraic equations and yields
  linear relationships between time derivatives of generalized
  coordinates/speeds (both independent and dependent) and independent
  generalized coordinates/speeds and exogenous inputs.  These linear ODE's
  effectively embed the constraints, and enable stability analysis and control
  system design without the need to explicitly address the constraints.  We
  numerically validate the results on a well studied model of a bicycle which
  includes both configuration and velocity constraints.
\keywords{symbolic dynamics, control, constrained multibody systems}
\end{abstract}

@@

%The system described by equations \ref{eq:configuration}-\ref{eq:dyndiffs} has
%$o-m$ velocity degrees of freedom.
Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
manifold embedded in a $n + o$ dimensional space, though conserved quantities
(e.g., total mechanical energy) may restrict trajectories to a lower
dimensional manifold.  In general this cannot be assumed, especially
in the presence of external control inputs $\bm{r}$ which may change the total
mechanical energy of the system.  While more than $p$ of the independent state
variables may be of interest (i.e., dependent coordinates or speeds may be of
interest in themselves), only $p$ quantities are truly independent.  Which of
these $p$ coordinates and speeds should be chosen as independent out of the
possible $n + o$ may be obvious for small systems where there is intuition
about the system behavior, but in general the optimal choice depends on how the
configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
hand, in systems with cyclic/ignorable coordinates (i.e., coordinates which do
not appear in the dynamic equations of motion), fewer than $p$ state
variables may be of interest; one example is the Whipple bicycle model where
$p=10$ but 5 of these state variables are cyclic coordinates (coordinates of
rear wheel contact point, heading, and both wheel angles) and may only be of
secondary interest.

@@

The choice on which state variables are taken to be independent may determine
whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
P_{ud}$ are nonsingular.  These matrices depend upon the configuration $\bm{q}$
and constant parameters; for certain configurations or parameters, it may be
that the choice of independent state variables cannot be arbitrary.  While some
systems may permit a choice of independent state variables which are valid for
all configurations of interest, others may not.  Methods for automatically
selecting the ``best'' choice of independent state variables are discussed in
\cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
constraint matrices to determine a set of indepedent speeds which will ensure
the nonsingularity of the aforementioned matrices.

@@

\begin{acknowledgements}
  This material is based upon work partially supported by the National Science
  Foundation under award 0928339 and two Google Summer of Code projects (2009,
  2011).  Jason Moore, Thomas Johnston, Evan Sperber, and Andrew Kickertz
  provided valuable feedback during discussions of multibody dynamics and
  control related issues.
\end{acknowledgements}

>>>>

\begin{abstract}
  We present a systematic procedure for linearizing equations of motion
  governing constrained multibody systems which have been derived with Kane's
  method.  The procedure avoids differential algebraic equations and yields
  linear relationships between time derivatives of generalized
  coordinates/speeds (both independent and dependent) and independent
  generalized coordinates/speeds and exogenous inputs.  These linear ODEs
  effectively embed the constraints, and enable stability analysis and control
  system design without the need to explicitly address the constraints.  We
  numerically validate the results on a well studied model of a bicycle which
  includes both configuration and velocity constraints.
\keywords{symbolic dynamics, control, constrained multibody systems}
\end{abstract}

@@

Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
manifold embedded in a $n + o$ dimensional space, though conserved quantities
(e.g., total mechanical energy) may restrict trajectories to a lower
dimensional manifold.  In general this cannot be assumed, especially
in the presence of external control inputs $\bm{r}$ which may change the total
mechanical energy of the system.  While more than $p$ of the independent state
variables may be of interest (i.e., dependent coordinates or speeds may be of
interest in themselves), only $p$ quantities are truly independent.  Which of
these $p$ coordinates and speeds should be chosen as independent out of the
possible $n + o$ may be obvious for small systems where there is intuition
about the system behavior, but in general the optimal choice depends on how the
configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
hand, in systems with cyclic/ignorable coordinates (i.e., coordinates which do
not appear in the dynamic equations of motion), fewer than $p$ state
variables may be of interest; one example is the Whipple bicycle model where
$p=10$ but 5 of these state variables are cyclic coordinates (coordinates of
rear wheel contact point, heading, and both wheel angles) and may only be of
secondary interest.

@@

The choice on which state variables are taken to be independent may determine
whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
P_{ud}$ are nonsingular.  These matrices depend upon the configuration $\bm{q}$
and constant parameters; for certain configurations or parameters, it may be
that the choice of independent state variables cannot be arbitrary.  While some
systems may permit a choice of independent state variables which are valid for
all configurations of interest, others may not.  Methods for automatically
selecting the ``best'' choice of independent state variables are discussed in
\cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
constraint matrices to determine a set of independent speeds which will ensure
the non-singularity of the aforementioned matrices.

@@

\begin{acknowledgements}
  This material is based upon work partially supported by the National Science
  Foundation under award 0928339 and two Google Summer of Code projects (2009,
  2011).  Jason Moore, Thomas Johnston, Evan Sperber, and Andrew Kickertz
  provided valuable feedback during discussions of multibody dynamics and
  control.
\end{acknowledgements}
